---
title: "HW 03"
author: "Cat Xia"
format:
  html:
    embed-resources: true
toc: true
execute:
  warning: false
  error: false
  message: false
---

## 1 - Du Bois challenge.

```{r loading packages}
#| label: loading packages

if (!require("pacman"))
  install.packages("pacman")

pacman::p_load(tidyverse, here, glue, scales, janitor, forcats, ggrepel, ggpubr, jpeg)

ggplot2::theme_set(ggplot2::theme_minimal())
options(width = 65)
knitr::opts_chunk$set(
  fig.width = 7,
  fig.asp = 0.618,
  fig.retina = 3,
  fig.align = "center",
  api = 300
)

```

```{r  Q1 load data}
#| label: Q1 load data

income <- read.csv(here("data", "income.csv"))
income <- clean_names(income)
glimpse(income)

paper_image <- readJPEG(here("images", "paper-4182513_1280.jpg"))
```

```{r du bois data wrangling}
#| label: du bois data wrangling

#pivots the table so that it creates two new columns, one for spending type and one for the amount
income_clean <- income|>
  pivot_longer(cols = -c(class, average_income), names_to = "spending_type", values_to = "amount")

#creates a order for the income to be in, fct_relevel consulted with documentation
income_clean$class <- fct_relevel(
  income_clean$class,
  "$1000 AND OVER", 
  "$750-1000", 
  "$500-750", 
  "$400-500", 
  "$300-400", 
  "$200-300", 
  "$100-200"
)

#adds the percent for the label on each bar
income_clean <- income_clean |>
  mutate("percent" = percent(amount/average_income, accuracy = 1))

#changes the order so it better adheres to the original graph
income_clean$spending_type <- fct_relevel(
  income_clean$spending_type,
  "other", "tax", "clothes", "food", "rent"
)
```

```{r du bois recreation}
#| label: du bois recreation

#creates a bar graph for each income level by amount, colored by spending
income_clean |>
  ggplot(aes(x = class, y = amount, fill = spending_type, order = spending_type)) +
  
 #Background image on the graph itself. Wasn't able to figure out how to get the background image to the whole graph
  background_image(paper_image) + #*
  
  #creates a bar graph where the y values are given by their identity
  geom_bar(stat = "identity", position = "stack") +
  
  #creates the percent labels
  geom_text(aes(label = percent), position = position_stack(vjust = 0.5), size = 4) +
  
  #flips the graph sideways
  coord_flip() +
  
  #colored with values closer to the original, though more saturated for ease of differentiation
  scale_fill_manual(values = c("grey","lightblue","pink", "#966FD6", "grey30")) +
  
  labs(
    title = "Income and Expenditure of 150 N*gro Families in Atlanta.GA.U.S.A",
    x = NULL,
    y = NULL,
    fill = "Anual Expenditure For"
  ) +
  theme(
    legend.position = "top",
    axis.ticks = element_blank(),
    axis.text.x = element_blank(),
    panel.grid = element_blank()
  ) 

```

\*<https://stackoverflow.com/questions/51255832/how-to-add-an-image-on-ggplot-background-not-the-panel>

Apologies, the graph is somewhat incomplete and not at all polished. I ran into a lot of bugs and now I am out of time.

## 2 - COVID survey - interpret

This graph is showing the the answers for each of the 6 questions regarding vaccines ("Vaccines are safe", "Vaccines make me feel safer at work", "Concern over Vaccine Side Effects", "Vaccine vetting process confidence", "vaccine information is trustworthy", and "I recommend the vaccine") for different identities from Age, Gender, Race, Ethnicity, Profession, or any prior vaccine experience for Covid and/or the flu. Starting from the top, for All data, most people tend to strongly agree on all questions regarding the vaccine except for "side effects" which is neutral. For age, there seems to be a trend where older people (\>30) are more likely to believe in the vaccine and are closer to 1. Surprisingly, there was not a huge difference in gender across the 4 different options, with the exception of Non-binary, where the vaccine was more likely to be believed as less safe but also more likely to not be as concerned about the side effects as male or female respondents. As race goes, Native Hawaiian/Pacific Islander and White respondents are more likely to have doubts about the vaccine. Both Hispanic and non-hispanic respondents were fairly equal. Surprisingly, nursing and medical professions had concerns over the side effects of the vaccine at a higher level than a good portion of the other population. Additionally, those who had had the vaccine were more wary of side effects, but those who haven't were much more generally doubtful, for both the COVID and Flu vaccine. Across the board, the question on vaccine safety and side effects garnered more varied responses than the others (much smaller error bars).

From my initial response, I agree with the graph in that those who haven't had vaccines this year are more likely to express higher levels of doubt around the covid vaccine for Flu and Covid. My intuition disagrees with the Nursing and Medical professions having significant doubts about the side effects, but more specifically how there are no other professions to compare to. I could be interpreting this response incorrect, and it may be something like "the profession that gives you the vaccine", although I am unsure. Lastly, I also thought that younger people from \<20 to 25 were more accepting of the vaccine and my intuition also disagrees with them having seemingly some of the greater concerns regarding the safety and side effects.

## 3 - COVID survey - reconstruct

```{r Q3 load data}
#| label: load data

covid <- read.csv(here("data", "covid-survey.csv"), skip = 1)
dim(covid)
```

```{r Q3 Data cleaning}
#| label: Q3 Data Cleaning

#cleans the row names
covid <- clean_names(covid)
covid <- covid |>
  #filters out the first row so that the second row can be the names of the col
  filter(if_any(-1, ~ !is.na(.))) #*
dim(covid)
```

-   \*Code after consulting with https://www.tidyverse.org/blog/2021/02/dplyr-1-0-4-if-any/. -1 from my own code, '\~ !is.na(.)' formatting consulted from AI after multiple bug issues.

```{r Q3 renaming variables}
#| label: renaming variables

covid_survey <- covid |>
  
  #changes all the col values in accordance with directions
  mutate(
    "exp_already_vax" = if_else(exp_already_vax == 0, "No", "Yes"),
    
    "exp_flu_vax" = if_else(exp_flu_vax == 0, "No", "Yes"),
    
    "exp_profession" = if_else(exp_profession == 0, "Medical", "Nursing"),
    
    "exp_gender" = case_when(
      exp_gender == 0 ~ "Male",
      exp_gender == 1 ~ "Female",
      exp_gender == 3 ~ "Non-binary third gender",
      exp_gender == 4 ~ "Prefer not to say",
    ),
    
    "exp_race" = case_when(
      exp_race == 1 ~ "American Indian/Alaskan Native",
      exp_race == 2 ~ "Asian",
      exp_race == 3 ~ "Black/African American",
      exp_race == 4 ~ "Native Hawaiian/Other Pacific Islander",
      exp_race == 5 ~ "White"
    ),
    
    "exp_ethnicity" = if_else(exp_ethnicity == 1, "Hispanic/Latino", "Non-Hispanic/Non-Latino"),
    
    #age changed to factor so it can be easily ordered.
    "exp_age_bin" = factor(case_when(
      exp_age_bin == 0 ~ "<20",
      exp_age_bin == 20 ~ "21-25",
      exp_age_bin == 25 ~ "26-30",
      exp_age_bin == 30 ~ ">30"),
      
      levels = c("<20", "21-25", "26-30", ">30")), #*
    
    #response changed to double to appear close to reference dataframe
    "response_id" = as.double(response_id)
  )

dim(covid_survey)
```

-   \*http://www.cookbook-r.com/Manipulating_data/Changing_the_order_of_levels_of_a_factor/

```{r pivot longer Q3 code given}
#| label: pivot longer Q3 code given

#creates a new column with the pivoted explanatory values. The row titles goes in explanatory, the row values go in explanatory values.
covid_survey_longer <- covid_survey |>
  pivot_longer(
    cols = starts_with("exp_"),
    names_to = "explanatory",
    values_to = "explanatory_value"
  ) |>
  
  #filters out NA values
  filter(!is.na(explanatory_value)) |>
  
  #creates a new column with all the response values. The row titles go in reponse, the values in response values.
  pivot_longer(
    cols = starts_with("resp_"),
    names_to = "response",
    values_to = "response_value"
  )
glimpse(covid_survey_longer)
```

The first pivot longer is taking all the explanatory values (ie background info) and putting it in explanatory_value while the second pivot longer is taking all the response and putting them under response_value.

```{r Q3 grouping}
#| label: Q3 grouping

#creates a tibble that has the mean, 10 quantile, and 90 quantile for each group and response.
covid_survey_summary_stats_by_group <- covid_survey_longer |>
  group_by(explanatory, explanatory_value, response) |>
  summarise(mean = mean(response_value, na.rm = TRUE),
            low = quantile(response_value, probs = 0.1, na.rm = TRUE), #*
            high = quantile(response_value, probs = 0.9, na.rm = TRUE))

#reorders the group so that it can be graphed nicely
covid_survey_summary_stats_by_group <- covid_survey_summary_stats_by_group[order(covid_survey_summary_stats_by_group$explanatory_value),] #**

glimpse(covid_survey_summary_stats_by_group)
```

\*https://www.tidyverse.org/blog/2020/03/dplyr-1-0-0-summarise/ \*\*https://bookdown.org/ndphillips/YaRrr/order-sorting-data.html

```{r Q3 summary not grouped}
#| label: summary not grouped Q3

#creates the mean, 10 quantile, 90 quantile, explanatory and explanatory placement values for all of the data
covid_survey_summary_stats_all <- covid_survey_longer |>
  group_by(response, .drop = FALSE) |>
  summarise(mean = mean(response_value, na.rm = TRUE),
            low = quantile(response_value, probs = 0.1, na.rm = TRUE), 
            high = quantile(response_value, probs = 0.9, na.rm = TRUE),
            explanatory = "ALL", #code inspired from discussion in HW channel
            explanatory_value = factor(""),
            .groups = "drop"
            ) 

covid_survey_summary_stats_all
```

```{r Q3 binding together two df}
#| label: binding together summary stats all and by group

#puts the two dataframes together
covid_survey_summary_stats <- rbind(covid_survey_summary_stats_all, covid_survey_summary_stats_by_group) #*

covid_survey_summary_stats
```

\*https://www.datacamp.com/doc/r/merging

```{r specifing the order of explanatory for easier faceting + facet titles}
#| label: specifying order of explanatory row for easier faceting + facet titles

#reordering explanatory so that it can be better graphed in order
covid_survey_summary_stats$explanatory <- factor(covid_survey_summary_stats$explanatory,
      levels = c("ALL", "exp_age_bin", "exp_gender", "exp_race", "exp_ethnicity", "exp_profession", "exp_already_vax", "exp_flu_vax"))

#to be labeled on the graph
questions <- c("resp_safety" = "Based on my understanding, I believe the vaccine is safe",
               "resp_feel_safe_at_work" = "Getting the vaccine will make me feel safer at work", 
               "resp_concern_safety" = "I am concerned about the safety and side effects of the vaccine",
               "resp_confidence_science" = "I am confident in the scientific vetting process for the new COVID vaccines",
               "resp_trust_info" = "I trust the information that I have recieved about the vaccine", 
               "resp_will_recommend" = "I will recommend the vaccine to family, friends, and community members")

#to be labeled on the graph
id_info <- c("ALL" = "All", 
             "exp_age_bin" = "Age", 
             "exp_gender" = "Gender", 
             "exp_race" = "Race", 
             "exp_ethnicity" = "Ethnicity", 
             "exp_profession" = "Profession",
             "exp_already_vax" = "Had COVID vaccine", 
             "exp_flu_vax" = "Had flu vaccine this year")
```

```{r Q3 recreating the graph}
#| label: recreating the graph for Q2


covid_survey_summary_stats |>
  #maps mean to explanatory value
  ggplot(aes(x = mean, y = explanatory_value)) +
  
  #creates the error bars for each point with the ends at the 10 and 90 quartile
  geom_errorbar(aes(xmin = low, xmax = high), width = 0.5) + #from documentation
  
  #creates the datapoint for each response
  geom_point(size = 1) +
  
  #facets based on reponse and explanatory
  facet_grid(cols = vars(response), 
             rows = vars(explanatory), 
             
             #"free_y" allows for the explanatory values on Y to be for the whole graph and not smushed in the facet of explanatory
             scales = "free_y", 
             space = "free_y",
             
             #takes the labels created above and maps them to each of the variables
             labeller = labeller(explanatory = as_labeller(id_info, default = label_wrap_gen(width = 10)),
                                 response = as_labeller(questions, default = label_wrap_gen(width = 14))
                                 )) +
  labs(
    x = "Mean Likert score \n (Error bars range from 10th to 90th percentile)",
    y = NULL
  ) +
  
  #attempt to better space out the y axis but not quite there
  scale_y_discrete(expand = c(0.1, 0.3)) + #*
  
  #theme with the grey boxes for facets
  theme_bw() +
  
  #changes the texts size and gets rid of all the grid elements
  theme(
    strip.text.x = element_text(size = 8, lineheight = 0.8),
    strip.text.y = element_text(size = 8, lineheight = 0.8, angle = 0),
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_text(size = 8),
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(size = 8)
  )
```

-   \*https://stackoverflow.com/questions/15400486/change-distance-between-x-axis-ticks-in-ggplot2

## 4 - COVID survey - re-reconstruct

```{r Q4 recreate summary stats}
#| label: Q4 recreate summary stats

#only change is that the low and high values are the 25 and 75th quantile respectively
covid_survey_summary_stats_by_group2 <- covid_survey_longer |>
  group_by(explanatory, explanatory_value, response) |>
  summarise(mean = mean(response_value, na.rm = TRUE),
            low = quantile(response_value, probs = 0.25, na.rm = TRUE), 
            high = quantile(response_value, probs = 0.75, na.rm = TRUE))

covid_survey_summary_stats_by_group2 <- covid_survey_summary_stats_by_group2[order(covid_survey_summary_stats_by_group2$explanatory_value),]

covid_survey_summary_stats_all2 <- covid_survey_longer |>
  group_by(response, .drop = FALSE) |>
  summarise(mean = mean(response_value, na.rm = TRUE),
            low = quantile(response_value, probs = 0.25, na.rm = TRUE), 
            high = quantile(response_value, probs = 0.75, na.rm = TRUE),
            explanatory = "ALL",
            explanatory_value = factor(""),
            .groups = "drop"
            ) 

covid_survey_summary_stats2 <- rbind(covid_survey_summary_stats_all2, covid_survey_summary_stats_by_group2)

covid_survey_summary_stats2$explanatory <- factor(covid_survey_summary_stats2$explanatory,
      levels = c("ALL", "exp_age_bin", "exp_gender", "exp_race", "exp_ethnicity", "exp_profession", "exp_already_vax", "exp_flu_vax"))

```

```{r Q4 graph with 25q and 75q}
#| label: Q4 graph with 25 quantile and 75 quantile

#same code as above but with the new dataframe with the 25th and 75th quantile values

covid_survey_summary_stats2 |>
  ggplot(aes(x = mean, y = explanatory_value)) +
  geom_errorbar(aes(xmin = low, xmax = high), width = 0.5) + #from documentation
  geom_point(size = 1) +
  facet_grid(cols = vars(response), 
             rows = vars(explanatory), 
             scales = "free_y", 
             space = "free_y",
             labeller = labeller(explanatory = as_labeller(id_info, default = label_wrap_gen(width = 10)),
                                 response = as_labeller(questions, default = label_wrap_gen(width = 14))
                                 )) +
  labs(
    x = "Mean Likert score \n (Error bars range from 10th to 90th percentile)",
    y = NULL
  ) +
  scale_y_discrete(expand = c(0.1, 0.3)) + #*
  theme_bw() +
  theme(
    strip.text.x = element_text(size = 8, lineheight = 0.8),
    strip.text.y = element_text(size = 8, lineheight = 0.8, angle = 0),
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_text(size = 8),
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(size = 8)
  )
```

The graph with 25th and 75th quartile error bars reaffirms that safety of the vaccine and vaccine side effects have the greatest diversity of answers, while the other ones tend to be much closer to 1 or 2, meaning people approved of the vaccine more in those areas. The differences in Age also seem to be diminished with the smaller quartiles as well as the differences in Nursing and Medical. The differences between having taken or not taken the COVID vaccine are still quite distinct, which makes sense, but the differences in Flu vaccine got smaller.

## 5 - COVID survey - another view

```{r Q5 clean data}

#creates a dataframe with the irrelevant rows removed
covid_response <- covid_survey[, c(9, 10, 11, 12, 13, 14)] #*

#pivots so the questions and scores are the only two columns
resp_only <- covid_response |>
  pivot_longer(
    cols = starts_with("resp"),
    names_to = "question",
    values_to = "score",
  )

#did not fix the NA values earlier so came back to fix them here
resp_only <- na.omit(resp_only) #**

```

\*https://www.datacamp.com/tutorial/subsets-in-r \*\*https://bookdown.org/rwnahhas/IntroToR/exclude-observations-with-missing-data.html

The Code chunk below and the diverging graph it generates was done following the guide made by https://rfortherestofus.com/2021/10/diverging-bar-chart

```{r Digerging graph data wrangling new variables}

#Creates a dataframe that calculates the number of occurances per score and the percent of the question it makes up
resp_only_sum <- resp_only |>
  
  #groups by each question and each score
  group_by(question, score) %>%
  
  #counts the occurances and stores in n_answers
  count(name = "n_answers") |>
  
  #then group by questions to get the total sum of each question
  group_by(question) |>
  
  #creates a new column that gives the percent that each score makes up
  mutate(percent_answers = n_answers/sum(n_answers)) |>
  ungroup() |>
  
  #creates a new column where the percent gets transformed into a percent value
  mutate(percent_answers_label = percent(percent_answers, accuracy = 1),
         score_fill = as.character(score))

#creates a new dataframe for the diverging that classifies the lower values as postive and the higher values as negative, with 3 being 0
resp_diverging <- resp_only_sum |>
  #creates the column where 3 and lower are pos, 4 and higher are negative
  mutate(percent_answers_neg = if_else(score <=3, percent_answers, -percent_answers)) |>
  #creates the label so that it is not negative
  mutate(percent_answers_label = percent(percent_answers, accuracy = 1))

glimpse(resp_diverging)
```

```{r divering graph, graphed}


resp_diverging|>
  #assigns x to each question anticipating that the coordinates will be flipped so that eventually it will be on the vertical
  ggplot(aes(x = question, y = percent_answers_neg, fill = score_fill)) +
  
  #creates the bar graph
  geom_col() +
  
  #unable to get the text below to work without being on top of each other
  geom_text(aes(label = percent_answers_label),
            position = position_stack(vjust = 0.5)) +
  
  #flips the graph
  coord_flip() +
  theme_minimal() +
  
  #assigns green for positive connotations and red for negative connotations.
  scale_fill_manual(values = c("#30c878", "#7CFC00", "skyblue", "#D2042D", "#953553")) +
  
  #wraps the text for each question for readability
  scale_x_discrete(labels = label_wrap(25)(questions)) +
  
  labs(
    title = "Percent response to COVID Vaccine Questions",
    y = "Score",
    x = NULL,
    fill = "Score Out of 5:") +
  
  #deletes all the unnecessary graph elements
  theme(
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_blank(),
    legend.position = "top"
  ) 
```

**A.** Image of a graph titled, "Percent response to COVID vaccine Questions that looks at the percent breakdown of scores (out of five, where one is strongly agree and 5 is strongly disagree) for six questions centered around the covid vaccine. The first question, "I will recommend the vaccine to family, friends, and community members", has 88% of respondents strongly agreeing, while the other responses were relatively equal. The second question, "I trust the information that I have recieved about the vaccine", has 77% of respondents strongly agreeing and 21% agreeing. After that, the disagree and neutral categories were slightly more than the disagree respondents, although the three make up very little of the percentage. The third question is "I am confident in the scientific vetting process for the new COVID vaccines" which has a 61% strongly agree, 14% agree, and a 17% strongly disagree, with the other two responses both being at 4%. The fourth question, "I am concerned about the safety and side effects of the vaccine" has 81% strongly agree, 10% agree, and minimal percent from the other responses. The fifth question, "Getting the vaccine will make me feel safer at work", has a 70% strongly agree, 23% agree, and minimal percentage for the other three. Lastly, the sixth question, "Based on my understanding, I believe the vaccine is safe", shows a significant deviation from the other questions, with 31% disagreeing, 24% agreeing, 22% strongly disagreeing, and only 12% strongly agreeing, leaving neutral responses at 11%. Ultimately, this graph highlights how most questions have respondents strongly agreeing, except for the question on vaccine safety, which is mixed, but tends slightly towards disagree.

```{r Not Diverging graph}
#code below is the same as above except it uses the prior dataframe before all the scores were made postive or negative
resp_only_sum|>
  ggplot(aes(x = question, y = percent_answers, fill = score_fill)) +
  geom_col() +
  
  #unable to get the text below to work without being on top of each other
  geom_text(aes(label = percent_answers_label),
            position = position_stack(vjust = 0.5)) +
  
  coord_flip() +
  theme_minimal() +
  scale_fill_manual(values = c("#30c878", "#7CFC00", "skyblue", "#D2042D", "#953553")) +
  scale_x_discrete(labels = label_wrap(25)(questions)) +
  labs(
    title = "Percent response to COVID Vaccine Questions",
    y = "Score",
    x = NULL,
    fill = "Score Out of 5:") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_blank(),
    legend.position = "top"
  ) 
```

**B.** Image of a graph titled, "Percent response to COVID vaccine Questions that looks at the percent breakdown of scores (out of five, where one is strongly agree and 5 is strongly disagree) for six questions centered around the covid vaccine. Unlike the previous graph, which is centered on neutral response, this graph is more concerned with the comparison of percentages for each score response per question. It highlights how each score differs and by how much, for example, strongly agreeing going from around 88% to 12% or agreeing varying between 7% and 24%. The rest of the values are the same for each question and those values are repeated from the ID above as follows: The first question, "I will recommend the vaccine to family, friends, and community members", has 88% of respondents strongly agreeing, while the other responses were relatively equal. The second question, "I trust the information that I have recieved about the vaccine", has 77% of respondents strongly agreeing and 21% agreeing. After that, the disagree and neutral categories were slightly more than the disagree respondents, although the three make up very little of the percentage. The third question is "I am confident in the scientific vetting process for the new COVID vaccines" which has a 61% strongly agree, 14% agree, and a 17% strongly disagree, with the other two responses both being at 4%. The fourth question, "I am concerned about the safety and side effects of the vaccine" has 81% strongly agree, 10% agree, and minimal percent from the other responses. The fifth question, "Getting the vaccine will make me feel safer at work", has a 70% strongly agree, 23% agree, and minimal percentage for the other three. Lastly, the sixth question, "Based on my understanding, I believe the vaccine is safe", shows a significant deviation from the other questions, with 31% disagreeing, 24% agreeing, 22% strongly disagreeing, and only 12% strongly agreeing, leaving neutral responses at 11%.

**C.** The graphs in a and b differ in that graph a is able to allow the view to better compare between responses (negative responses together and positive responses together). On the other hand, graph b allows the viewer to compare the percent makeup of each answer and which ones score more positive overall, and if not, how they break down.
